---
title: "20190916"
output: html_notebook
---

## 4.4 For-loops

The formula for the sum of the series 1+2+...+n is n(n+1)/2. Using what we learned about functions we can creat one that computes the Sn:
```{r}
compute_s_n <- function(n){
  x<-1:n
  sum(x)
}
```

n=1,...,25일때 Sm을 어떻게 계산할까? 이 때 for-loops를 사용한다.
```{r}
for(i in 1:5){
  print(i)
}
```

our Sn example:
```{r}
m <- 25
s_n <- vector(length = m)
for(n in 1:m){
  s_n[n] <- compute_s_n(n)
}
```

```{r}
n <- 1:m
plot(n,s_n)
```

```{r}
head(data.frame(s_n = s_n, formula = n*(n+1)/2))
```

```{r}
plot(n, s_n)
lines(n,n*(n+1)/2)
```

## 4.5 Vectorization and functionals

*Vectorization* is preferred over for-loop since it results in shorter and clearer code. A vectorized function is a function that will apply the same operation on each of the vectors.
```{r}
x <- 1:10
sqrt(x)
y <- 1:10
x*y
```

This time, we don't need for-loops.The function we just wrote,`compute_s_n`, does not work on each entry of `n`: 
```{r}
n <- 1:25
compute_s_n(n)
```

Functionals are functions which can apply the same function to each entry in all type of data. 
For numeric, logical, and character vector: `sapply`

The funtion `sapply` can do element-wise operation.

```{r}
x <- 1:10
sapply(x,sqrt)
```
This result is a vector of same length as the original `x`. 

The for-loop above can be written:
```{r}
n <- 1:25
s_n <- sapply(n,compute_s_n)
plot(n,s_n)
```
Other functials are `apply`, `lapply`, `tapply`, `mapply`, `vapply`, and `replicate`. 

## 4.6 Exercises

1. What will this conditional expression return?
```{r}
x <- c(1,2,-3,4)

if(all(x>0)){
  print("All Postives")
} else{
  print("Not all positives")
}
```

2. Which of the following expressions is always `FALSE` when at least one entry of a logical vector `x` is TRUE?

```{r}
x <- c(TRUE, TRUE, TRUE)

all(!x)
```

A. all(x)

B. any(x)

C. any(!x)

**D. all(!x)**

3. The function `nchar` tells you how many characters long a character vector is.

Write a line of code that assigns to the object `new_names` the state abbreviation when the state name is longer than 8 characters.
```{r}
dslabs::murders
data(murders)
new_names <- ifelse(((nchar(murders$state)) > 8), murders$abb, murders$state)
new_names
```

4. Create a function `sum_n` that for any given value, say n, computes the sum of the integers from 1 to n (inclusive). Use the function to determine the sum of integers from 1 to 5,000.
```{r}
sum_n <- function(n){
  sum(1:n)
}
sum_n(5000)
```

5. Create a function `altman_plot` that takes two arguments, `x` and `y`, and plots the difference against the sum.
```{r}
altman_plot <- function(x,y){
  a = x-y
  b = x+y
  plot(x,y)
}
altman_plot(x,y)
```
??? 해당 plot이 무엇을 보고자 하는지 조차 이해되지 않았으며 function을 지정했지만 length가 달라 실행되지 않는다.


6. After running the code below, what is the value of `x`?
```{r}
x <- 3
my_func <- function(y){
  x <- 5
  y+5
}
x
```

7. Write a function `compute_s_n` that for any given n computes the sum Sn = 1^2 + 2^2 + 3^2 +...n^2.Report the value of the sum when  n=10.
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}
compute_s_n(10)
```

8. Define an empty numerical vector `s_n` of size 25 using 
`s_n <- vector("numeric", 25)` and store in the results of S1,S2,...S25 using a for-loop.
```{r}
m <- 25
s_n <- vector("numeric", 25)
for(n in 1:m){
  s_n[n] <- compute_s_n(n)
}
```


9. Repeat exercise 8, but this time use `sapply`.
```{r}
n <- 1:25
sapply(n,compute_s_n)
```

10. Repeat exercise 8, but this time use `map_dbl.`
```{r}
library(purrr)
map_dbl(n,compute_s_n)
```

11. Plot Sn versus n. Use points defined by n = 1,...,25. 
```{r}
n <- 1:25
plot(n,s_n)
lines(n,s_n)
```

12. Confirm that the formular for this sum is Sn =n(n+1)(2n+1)/6
```{r}
new_func <- function(n){
  p <- n*(n+1)*(2*n+1)/6
  p
}
sapply(n,compute_s_n)
sapply(n,new_func) 
```
logical적으로 일치함을 TRUE로 나타낼 수 있다면 좋을 것 같지만 성공하지 못했다. 

This is the last of chapter 4.






