---
title: "R Notebook"
output: html_notebook
---

---
title: "20190911_141_choi"
output: html_notebook
---

# Chapter 4 Programming basics
In this section, we introduce three key programming concepts: 1) conditional expressions, 2)for-loops and 3)functions.

## 4.1 Conditional expressions

Conditional expressions are used for *flow control*. The most common conditional expression is the if-else statement. 
The basic idea is to print the reciprocal of `a` inless `a` is 0:

```{r}
a <- 0

if(a!=0){
  print(1/a)
} else{
    print("No reciprocal for 0.")}
```

```{r}
library(dslabs)
data("murders")
murder_rate <- murders$total/murders$population *100000
```

Which states have a murder rate lower than 0.5 per 100,000? The `if` statement protects us from the case in which no state satisfies the condition.

```{r}
ind <- which.min(murder_rate)

if(murder_rate[ind] <0.5){
  print(murders$state[ind])
} else{
  print("No state has murder rate that low")
}
```

If we try it again with a rate of 0.25, we get a different answer:
```{r}
if(murder_rate[ind] < 0.25){
  print(murders$state[ind]) 
} else{
  print("No state has a murder rate that low.")
}
```

The function `ifelse` takes three arguments: a logical and two possible answers. If the logical is `TRUE`, the value in the second argument is returned and if `FALSE`, the value in the third argument is returned.
```{r}
a<- 0
ifelse(a>0, 1/a, NA)
```

```{r}
data(na_example)
no_nas<- ifelse(is.na(na_example),0,na_example)
sum(is.na(no_nas))
```

`any` function은 어떤 entries가 `TRUE`일때 `TRUE`를 보내고 `all` function은 모든 entries가 'TRUE'일때 'TRUE'을 보낸다.

```{r}
z <- c(TRUE, TRUE, FALSE)
any(z)
all(z)
```

## 4.2 Defining functions

We can compute the average of a vector x using the `sum` and `length` functions: `sum(x)/length(x)`.

```{r}
avg <- function(x){
  s <- sum(x)
  n <- length(x)
  s/n
}
```
Now `avg` is a function that computes the mean:

```{r}
x <- 1:100
identical(mean(x), avg(x))
```

Variables defined inside a function are not saved in the workspace.
```{r}
s <- 3
avg(1:10)
s
```

The function `function` tells R you are about to define a function. 

my_function <- function(VARIABLE_NAME){
  perform operations on VARIABLE_NAME and calculate VALUE
  VALUE
}
```{r}
avg <- function(x, arithmetic = TRUE){
  n<- length(x)
  ifelse(arithmetic, sum(x)/n, prod(x)^(1/n))
}
```


## 4.3 Namespaces

여러가지 packages를 사용할 때 두 개의 다른 fucnion이 같은 이름을 사용하는 경우가 있다. 

Both *dplyr* and the R-base *stats* package define a filter function.

The following objects are masked from ‘package:stats’:

    filter, lag

The following objects are masked from ‘package:base’:

    intersect, setdiff, setequal, union

When we type `filter`, it uses the *dplyr* one. But what if we want to use the *stats* version?

Use different namespaces. 
```{r}
search()
```
dplyr filter대신 stats filter을 사용하고싶은데 search list에서 dplyr이 첫번째라면 double olons(`::`):
```{r}
stats::filter
#if we want to use dplyr filter
dplyr::filter
```

