---
title: "rbasic_20190906_141_choi"
output: html_notebook
---
## 3.2 The very basics 

### 3.2.1 Objects

Quadratic fomula를 이용해 방정식의 해를 구할 때 values의 값(a, b, c)에 따라 변한다. Programming languages의 장점은 define variables and write expressions with these variables이다. 
To solve X^2+X-1=0, then we define:
```{r}
a <- 1
b<- 1
c <- -1
```

which stores the values for later use. 
We use `<-` to assign values to the variables. `=`도 가능하지만 `<-`을 추천한다.

Copy and paste the code above into your console to define the three variables. 
아무것도 뜨지 않으면 objects가 성공적으로 defined되었지만 실수를 했다면 error message가 뜰 것이다.

To see the value stored in a variable, evaluate `a`.
```{r}
a
```

using print is also okay.
```{r}
print(a)
```

"object"는 stuff that is stored in R을 의미한다. 

### 3.2.2 The workspace

You can see all the variables saved in your workspace by typing:
```{r}
ls()
```

If you try to recover the value of a variable that is not in your workspace, you receive an error.

values가 variables에 저장되었으니 equation의 해를 구하기위해 근의 공식을 사용하면,

```{r}
(-b+sqrt(b^2-4*a*c))/(2*a)
(-b-sqrt(b^2-4*a*c))/(2*a)
```

### 3.2.3 Funtions

Variables를 define하고나서는 data analysis process는 data에 적용되는 series of `function`이다. 

`install.packages`, `library`, `ls`그리고 `sqrt`는 사용해보았다. 아직 define되지 않았기 때문에 workspace에서 보이지 않지만 바로 사용 가능하다. 

In general, we need to use parentheses to evaluate a funtion. `ls`와 `ls()`는 다르게 작동한다. 

`log`라는 funtion을 사용해보자.
```{r}
log(8)
log(a)
```

Funtion을 알고싶다면 `help`를 사용한다.
```{r}
help("log")
```

shorthand를 사용가능하다.
```{r}
?log
```

`log` needs `x` and `base` to run. Help를 이용해서 required argument와 optional argument를 확인할 수 있다. 

The base of the funtion `log` defalut to `base=exp(1)` making `log` the natural log by defalt.
```{r}
args(log)
```

change default values;
```{r}
log(8, base = 2)
```

x를 사용하면;
```{r}
log(x=8, base = 2)
```

argument를 사용하지 않으면 자동으로 적용된다.
```{r}
log(8,2)
```

argument의 순서는 상관없다.
```{r}
log(base=2, x=8)
```

argument는 `<-`가 아닌 `=`을 사용해야 한다.

Funtions need the parenteses라는 rule의 exception이 있다.
```{r}
2^3
```

```{r}
help("+")
?"+"
help(">")
?">"
```

### 3.2.4 Other

You can see all the available datasets by typing:
```{r}
data()
```

This shows you the object name for these datasets. These can be used by simply typing the name.
```{r}
co2
```

R will show you Mauna Loa atmospheric CO2 concentration data.

Mathematical quantities-constant `pi` and `infinite`
```{r}
pi
Inf+1
```

### 3.2.5 Variable names

Variable name은 어떤 leter도 될 수 있다. 하지만 variable names have to start with a letter, can't contain spaces an should not be variables that are predefined in R.

`install.packages`를 `install.packages <- 2`로 쓰지 마라.

A nice convention to follow is to use meaningful words that describe what is stored, use only lower case, and use underscores as a substitute for spaces

```{r}
solution_1 <- (-b + sqrt(b^2 - 4*a*c)) / (2*a)
solution_2 <- (-b - sqrt(b^2 - 4*a*c)) / (2*a)
```

#3.2.6 Saving your workspace

Values remain in the workspace until you end your session or erase them with the function `rm`. 

we recommend you assign the workspace a specific name. Function `save` or `save.image`. To load, use the funtion `load`. When saving a workspace, we recommend the suffix `rda` or `rData`. 

### 3.2.7 Motivating scripts

To solve another equation such as 3x^2+2x−1, redefine the variables and recompute the solution:
```{r}
a<-3
b<-2
c<--1
(-b + sqrt(b^2 - 4*a*c)) / (2*a)
(-b - sqrt(b^2 - 4*a*c)) / (2*a)
```

### 3.2.8 Commenting your code

만약 #으로 R code를 시작하면 it is not evaluated.

```{r}
## code to compute solution to quadratic equation of the form ax^2+bx+c
## define the variables
a<- 3
b<- 2
c<- -1

## now compute the solution
(-b + sqrt(b^2-4*a*c))/(2*a)
(-b - sqrt(b^2-4*a*c))/(2*a)
```

## 3.3 Exercises

1. What is the sum of the first 100 positive integers? The formula for the sum of integers 1 through n is  
n(n+1)/2. Define n = 100 and then use R to compute the sum of 1 through 100 using the formula. What is the sum?
```{r}
n <- 100
n*(n+1)/2
```

2. Now use the same formula to compute the sum of the integers from 1 through 1,000.
```{r}
n <- 1000
n*(n+1)/2
```


3. Look at the result of typing the following code into R:
```{r}
n <- 1000
x <- seq(1, n)
sum(x)
```

Based on the result, what do you think the functions seq and sum do? You can use the `help` system:

B. `seq` creates a list of numbers and `sum` adds them up.


4. In math and programming, we say that we evaluate a function when we replace the argument with a given number. So if we type `sqrt(4)`, we evaluate the `sqrt` function. In R, you can evaluate a function inside another function. The evaluations happen from the inside out. Use one line of code to compute the log, in base 10, of the square root of 100.
```{r}
log(base=10, sqrt(100))
```

5. Which of the following will always return the numeric value stored in `x`? You can try out examples and use the help system if you want.

B. `log10(x^10)`


## 3.4 Data types

The funtion `class` helps us determine what type of object we have:
```{r}
a <- 2
class(a)
```

## 3.5 Data frames

Up to now, the variables we have defined are just one number. The most common way of storing a dataset in R is in a data frame. Data frame is a table with rows representing observations and the different variables reported for each observation defining the columns.
Data frames은 dataset에 유용하다.

Data analysis challenges start with data stored in a data frame. You can access this dataset by loading the **dslabs** library and loading the `murders` dataset using the `data` function:
```{r}
install.packages("dslabs")
library(dslabs)
data(murders)
class(murders)
```

### 3.5.1 Examining an object

The funtion `str` is useful for finding out more about the structure of an object:

```{r}
str(murders)
```

We see that the table has 51 rows (50 states plus DC) and five variables. We can show the first six lnes using the function `head`:
```{r}
head(murders)
```
In this dataset, each state is considered an observation and five variables are reported for each state.

### 3.5.2 The accessor: `$`

To access the different variables represented by columns included in this data frame, use accessor operator `$`.
```{r}
murders$population
```

We can quickly access the variable names using:
```{r}
names(murders)
```

the order of entries in `murder$population` preserves the order of the rows in our data table.

### 3.5.3 Vectors: numerics, characters, and logical

The object `murders$population` is not one number but several. We call these types of objects vectors. A single number is technically a vector of length 1, but in general we use the term vectors to refer to objects with several entries. The funtion `length` tells you how many entries are in the vector:

```{r}
pop <- murders$population
length(pop)
```

This particualar vector is numeric and every entry must be a number.
```{r}
class(pop)
```

To store character strings, we use vector of class *character* and all eneries need to be a character.
```{r}
class(murders$state)
```

*Logical vectors* must be either `TRUE` or `FALSE`. 
```{r}
Z <- 3 == 2
Z
class(Z)
```
 `==` is a relational operator asking if 3 is equal to 2. 
 `=` is used when you actualy assign a variable, but `==` is used for equality.
 
```{r}
?Comparison
```
 
### 3.5.4  Factors

In the `murder` dataset, region을 character vector라고 생각할 수 있지만 *factor*이다. Factors are useful for storing categorical data. 
```{r}
class(murders$region)
```

We can see that there are only 4 regions by using the `levels` function:

```{r}
levels(murders$region)
```

R stores these level as integres and keeps a map to keep track of the labels. This is more memory efficient than storing all the characters.

Levels은 factor object에서 보이는 순서와 다른 순서를 갖고 있다. The default is for the levels to follow alphabetical order. The funtion `reorder` lets us change the order of the levels of a factor variable based on a summary computed on a vector.

만약 우리가 alphabetical order이 아니라 total number of murders의 levels of the region을 원한다면 `reorder`을 사용하고 specifiy a data sumary to dtermine the order. The following code takes the sum of the total murders in each region, the reorders the factor following these sums.

```{r}
region <- murders$region
value <- murders$total
region <- reorder(region, value, fun = sum)
levels (region)
```
새로운 순서는 Northeast가 murders이 제일 적고 South가 제일 많음을 의미한다. 


### 3.5.5 Lists

Data frames are a special case of *lists*. They are useful because you can store any combination of different types. 

### 3.5.6 Matrices

Matrices는 data frames과 비슷하다:they have rows and columns. However, like numeric, character and logical vectors, entries는 한종류이다. 
`matrix` funtion을 이용해 matrix를 define할 수 있다.
```{r}
mat <- matrix(1:12,4, 3)
mat
```

You can access specific entries in a matrix using square brackets(`[`). Second row, third column:
```{r}
mat[2,3]
```

Entire second row:
```{r}
mat[2, ]
```

```{r}
mat[ ,2:3]
```

```{r}
murders[25,1]
murders[2:3, ]
```

## 3.6 Exercises

1. Load the US murders dataset.
```{r}
library(dslabs)
data(murders)
str(murders)
```

Use the function `str` to examine the structure of the murders object. We can see that this object is a data frame with 51 rows and five columns. Which of the following best describes the variables represented in this data frame?

C. The state name, the abbreviation of the state name, the state’s region, and the state’s population and total number of murders for 2010.


2. What are the column names used by the data frame for these five variables?

```{r}
names(murders)
```

3. Use the accessor $ to extract the state abbreviations and assign them to the object a. What is the class of this object?

```{r}
murders$abb
a <- murders$abb
class(a)
```

4. Now use the square brackets to extract the state abbreviations and assign them to the object b. Use the identical function to determine if a and b are the same.
```{r}
data("murders")
b <- murders[ ,2]
a == b
```


5. We saw that the region column stores a factor. You can corroborate this by typing:

```{r}
class(murders$region)
```

With one line of code, use the function levels and length to determine the number of regions defined by this dataset.
```{r}
levels(murders$region)
length(murders$region)
```


6. The function `table` takes a vector and returns the frequency of each element. You can quickly see how many states are in each region by applying this function. Use this function in one line of code to create a table of states per region.

```{r}
table(murders$region)
```






