---
title: "20191014"
output: html_notebook
---

# Chapter 12 Robust summaries

```{r}
library(tidyverse)
library(dslabs)
```

## 12.1 Outliers

Outliers are distinguishable value that out of the expected range, usually too big or too samll.

Suppose a colleague is charged with collecting demography data for a group of males. The data report height in feet and are stored in the object:
```{r}
library(dslabs)
data("outlier_example")
str(outlier_example)
```
= approximated by a normal distribution

Summarize with avg and sd:
```{r}
mean(outlier_example)
sd(outlier_example)
```
this group of males is much taller than usual!

Unexpected: SD is over 7 inches.

A quick plot reveals the problem:
```{r}
qplot(outlier_example)
```
at least one value that is nonsensical
= a height of 180 feet is impossible. 

The boxplot detects this point as an outlier:
```{r}
boxplot(outlier_example)
```


## 12.2 Median
If we have outlier, the average can become very large.

The median is robust to such outliers.

With this data the median is:
```{r}
median(outlier_example)
```
The median is what boxplots display as a horizontal line.


## 12.3 The inter quartile range (IQR)

The box in boxplots are defined by the first and third quartile
: the variability in the data: 50% of the data is within this range.

Inter quartile range (IQR)
: The difference between the 3rd and 1st quartile (or 75th and 25th percentiles) 
: robust to outliers (large values do not affect it)

IQR/1.349 = normally distributed data

We can see that this works well in our example since we get a standard deviation estimate of:
```{r}
IQR(outlier_example)/1.349
```
which is about 3 inches.


## 12.4 Tukey's definition of an outliers
The top whisker ends at the 75th percentile plus 1.5 × IQR. 
Similarly the bottom whisker ends at the 25th percentile minus 1.5 × IQR. 

If we define the first and third quartiles as  Q1 and  Q3 respectively, then an outlier is anything outside the range:
[Q1-1.5 x(Q3-Q1),Q3+1.5x(Q3-Q1)]

When the data is normally distributed, the standard units of these values are:
```{r}
q3 <- qnorm(0.75)
q1 <- qnorm(0.25)
iqr <- q3 - q1
r <- c(q1 - 1.5*iqr, q3 + 1.5*iqr)
r
```
Using the `pnorm` function, we see that 99.3% of the data falls in this interval.

If we want an outlier to be rarer, we can increase the 1.5 to a larger number. Tukey also used 3 and called these far out outliers. With a normal distribution, 100% of the data falls in this interval.

In the `geom_boxplot` function, this can be controlled by the `outlier.size` argument, which defaults to 1.5.

The 180 inches measurement is well beyond the range of the height data:
```{r}
max_height <- quantile(outlier_example, 0.75) + 3*IQR(outlier_example)
max_height
```

If we take this value out, we can see that the data is in fact normally distributed as expected:
```{r}
x <- outlier_example[outlier_example < max_height]
qqnorm(x)
qqline(x)
```

## 12.5 Median absolute deviation
Outlier가 있을 때 robustly standard deviation을 할고 싶으면 median absolute deviation (MAD) 를 사용하면 된다.
먼저 median을 계산하고 각 valuedml와 median과의 거리를 계산하면 된다.
MAD = median of the distances between the value and the median

The `mad` function is used for this.

For the height data, we get a MAD of:
```{r}
mad(outlier_example)
```
which is about 3 inches.


## 12.6 Exercises
We are going to use the HistData package. If it is not installed you can install it like this:
```{r}
#install.packages("HistData")
```

Load the height data set and create a vector `x` with just the male heights used in Galton’s data on the heights of parents and their children from his historic research on heredity.
```{r}
library(HistData)
data(Galton)
x <- Galton$child
```

1. Compute the average and median of these data.
```{r}
mean(x)
median(x)
```

2. Compute the median and median absolute deviation of these data.
```{r}
median(x)
mad(x)
```

3. Now suppose Galton made a mistake when entering the first value and forgot to use the decimal point. You can imitate this error by typing:
```{r}
x_with_error <- x
x_with_error[1] <- x_with_error[1]*10
```
How many inches does the average grow after this mistake?
```{r}
mean(x_with_error)
mean(x_with_error) - mean(x) 
```

4. How many inches does the SD grow after this mistake?
```{r}
sd(x_with_error) - sd(x)
```

5. How many inches does the median grow after this mistake?
```{r}
median(x_with_error) - median(x)
```

6. How many inches does the MAD grow after this mistake?
```{r}
mad(x_with_error) - mad(x)
```

7. How could you use exploratory data analysis to detect that an error was made?
*C. A boxplot, histogram, or qq-plot would reveal a clear outlier.*

8. How much can the average accidentally grow with mistakes like this? Write a function called `error_avg` that takes a value `k` and returns the average of the vector `x` after the first entry changed to `k`. Show the results for `k=10000` and `k=-10000`.
```{r}
error_avg <- function(k){
  x[1] <- k
  return(mean(x))
}
error_avg(10000)
error_avg(-10000)
```


## 12.7 Case study: self-reported study

The heights we have been looking at are not the original hieghts reported by students. The original reported heights are also included in the dslabs package and can be loaded like this:
```{r}
library(dslabs)
data("reported_heights")
head(reported_heights)
```

Height is a `character vector` so we create a new column with the `numeric version`:
```{r}
reported_heights <- reported_heights %>%
  mutate(original_heights = height, height = as.numeric(height))
```
Note that we get a warning about NAs. 

This is because some of the self reported heights were not numbers. We can see why we get these:
```{r}
reported_heights %>% 
  filter(is.na(height)) %>% 
  head()
```

Some students self-reported their heights using feet and inches rather than just inches. Others used centimeters and others were just trolling. For now we will remove these entries:
```{r}
reported_heights <- filter(reported_heights, !is.na(height))
```

If we compute the average and standard deviation, we notice that we obtain strange results. The average and standard deviation are different from the median and MAD:
```{r}
reported_heights %>% 
  group_by(sex) %>%
  summarize(average = mean(height), sd = sd(height),
            median = median(height), MAD = mad(height))
```

This suggests that we have outliers, which is confirmed by creating a boxplot:
```{r}
reported_heights %>% 
  group_by(sex) %>%
  ggplot(aes(sex, height)) +
  geom_boxplot()
```

We can see some rather extreme values. To see what these values are, we can quickly look at the largest values using the arrange function:
```{r}
reported_heights %>% 
  arrange(desc(height)) %>% 
  top_n(10, height)
```
The first seven entries look like strange errors. However, the next few look like they were entered as centimeters instead of inches. Since 184 cm is equivalent to six feet tall, we suspect that 184 was actually meant to be 72 inches.

We can review all the nonsensical answers by looking at the data considered to be far out by Tukey:
```{r}
max_height <- quantile(reported_heights$height, .75) + 3*IQR(reported_heights$height)
min_height <- quantile(reported_heights$height, .25) - 3*IQR(reported_heights$height)
c(min_height, max_height)
```

```{r}
reported_heights %>% 
  filter(!between(height, min_height, max_height)) %>% 
  select(original_heights) %>%
  head(n=10)
```
we see two common mistakes: 
1) entries in centimeters turn out to be too large
2) entries of the form x.y (x = feet,y = inches) turn out to be too small.

Some of the even smaller values, such as 1.6, could be entries in meters.







