---
title: "20190923"
output: html_notebook
---

# 5.7 Summarizing data
In this section, we cover two new dplyr verbs that make these computations easier: `summarize` and `group_by`. We learn to access resulting values using the `pull` function.

## 5.7.1 `summarize`
The `summarize` function compute summary statistics with intuitive and readable code.
The `heights` dataset includes heights and sex reported by students in an in-class survey.

```{r}
library(dplyr)
library(dslabs)
data(heights)
```

The following code computes the average and standard deviation for females:
```{r}
s <- heights %>% 
  filter(sex == "Female") %>%
  summarize(average = mean(height), standard_deviation = sd(height))
s
```
We use`average` and `standard_deviation`. Because the resulting table stored in `s` is a data frame, we can access the components with the accessor `$`:
```{r}
s$average
s$standard_deviation
```

`summarize` is aware of the variable names and we can use them directly. We can compute any other summary that operates on vectors and returns a single value. 

For example, we can add the median, min and max heights like this:
```{r}
heights %>% 
  filter(sex == "Female") %>%
  summarize(median = median(height), minimum = min(height), maximum = max(height))
```

We can obtain these three values with just one line using the `quantile` function: for example, `quantile(x, c(0,0.5,1))` returns the min (0th percentile), median (50th percentile), and max (100th percentile) of the vector x.
```{r}
heights %>% 
  filter(sex == "Female") %>%
  summarize(range = quantile(height, c(0, 0.5, 1)))
```
With the function `summarize`, we can only call functions that return a single value. 

let’s compute the average murder rate for the United States.
```{r}
data(murders)
murders <- murders %>% mutate(rate = total/population*100000)
```
Remember that the US murder rate is not the average of the state murder rates:
```{r}
summarize(murders, mean(rate))
```
The US murder rate is the total number of murders in the US divided by the total US population. So the correct computation is:
```{r}
us_murder_rate <- murders %>%
  summarize(rate = sum(total) / sum(population) * 100000)
us_murder_rate
```

## 5.7.2 `pull`
The `us_murder_rate` object defined above represents just one number. Yet we are storing it in a data frame:
```{r}
class(us_murder_rate)
```

when a data object is piped that object and its columns can be accessed using the `pull` function. To understand what we mean take a look at this line of code:
```{r}
us_murder_rate %>% pull(rate)
```
This returns the value in the `rate` column of` us_murder_rate` making it equivalent to `us_murder_rate$rate.`

To get a number from the original data table with one line of code we can type:
```{r}
us_murder_rate <- murders %>% 
  summarize(rate = sum(total) / sum(population) * 100000) %>%
  pull(rate)

us_murder_rate
class(us_murder_rate)
```
which is now a numeric.

## 3.7.3 Group then summarize with `group_by`
we may want to compute the average and standard deviation for men’s and women’s heights separately. The `group_by` function helps us do this.

If we type this:
```{r}
heights %>% group_by(sex)
```
This is now a special data frame called a grouped data frame and dplyr functions, in particular `summarize`, will behave differently when acting on this object.

When we summarize the data after grouping, this is what happens:
```{r}
heights %>% 
  group_by(sex) %>%
  summarize(average = mean(height), standard_deviation = sd(height))
```

The summarize function applies the summarization to each group separately.

For another example, let’s compute the median murder rate in the four regions of the country:
```{r}
murders %>% 
  group_by(region) %>%
  summarize(median_rate = median(rate))
```

# 5.8 Sorting data frames
We know about the `order` and `sort` function, but for ordering entire tables, the dplyr function `arrange` is useful. For example, here we order the states by population size:
```{r}
murders %>%
  arrange(population) %>%
  head()
```
With `arrange` we get to decide which column to sort by. To see the states by population, from smallest to largest, we arrange by `rate` instead:
```{r}
murders %>% 
  arrange(rate) %>% 
  head()
```
= ascending order

In dplyr, the function `desc` transforms a vector so that it is in descending order. 
To sort the table in descending order, we can type:
```{r}
murders %>% 
  arrange(desc(rate)) %>% 
  head()
```

## 5.8.1 Nested sorting
A third column can be used to break ties between first and second and so on. 
Here we order by `region`, then within region we order by murder rate:
```{r}
murders %>% 
  arrange(region, rate) %>% 
  head()
```
첫번째와 두번째의 기준을 모두 적용시켜서 arrange시킨다.

## 5.8.2 The top n
지금까지 `head`를 사용했지만 더 larger proportion을 보고싶다면 `top_n`을 사용할 수 있다.
**top_n(data frame, number of row, the variable to filter)**
Here is an example of how to see the top 10 rows:
```{r}
murders %>% top_n(10, rate)
```
Rows are not sorted by `rate`, only filtered. If want to sort, we need to use `arrange`. 
Note that if the third argument is left blank, `top_n`, filters by the last column.


# 5.9 Exercises
For these exercises, we will be using the data from the survey collected by the United States National Center for Health Statistics (NCHS). This center has conducted a series of health and nutrition surveys since the 1960’s. Starting in 1999, about 5,000 individuals of all ages have been interviewed every year and they complete the health examination component of the survey. Part of the data is made available via the NHANES package. Once you install the NHANES package, you can load the data like this:
```{r}
library(NHANES)
data(NHANES)
```
The NHANES data has many missing values. Remember that the main summarization function in R will return `NA` if any of the entries of the input vector is an `NA`. Here is an example:
```{r}
library(dslabs)
data(na_example)
mean(na_example)
sd(na_example)
```
To ignore the `NA`s we can use the `na.rm` argument:
```{r}
mean(na_example, na.rm = TRUE)
sd(na_example, na.rm = TRUE)
```

1. We will provide some basic facts about blood pressure. First let’s select a group to set the standard. We will use 20-29 year old females. `AgeDecade` is a categorical variable with these ages. Note that the category is coded like " 20-29“, with a space in front! What is the average and standard deviation of systolic blood pressure as saved in the BPSysAve variable? Save it to a variable called `ref`.

Hint: Use `filter` and `summarize` and use the `na.rm = TRUE` argument when computing the average and standard deviation. You can also filter the NA values using `filter.`
```{r}
ref <- NHANES %>% filter(Gender == "female") %>% filter(AgeDecade == " 20-29") %>%
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
ref
```


2. Using a pipe, assign the average to a numeric variable `ref_avg.` Hint: Use the code similar to above and then `pull.`
```{r}
ref_avg <- ref$average
ref_avg
```
```{r}
ref_avg <- ref %>% pull(average)
ref_avg
```


3. Now report the min and max values for the same group.
```{r}
NHANES %>% filter(Gender == "female") %>% filter(AgeDecade == " 20-29") %>%
  summarize(maximum = max(BPSysAve, na.rm = TRUE), minimum = min(BPSysAve, na.rm = TRUE))
```

4. Compute the average and standard deviation for females, but for each age group separately rather than a selected decade as in question 1. Note that the age groups are defined by `AgeDecade.` Hint: rather than filtering by age and `gender`, filter by Gender and then use `group_by.`
```{r}
NHANES %>% filter(Gender == "female") %>% group_by(AgeDecade) %>%
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

5. Repeat exercise 4 for males.
```{r}
NHANES %>% filter(Gender == "male") %>% group_by(AgeDecade) %>%
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

6. We can actually combine both summaries for exercises 4 and 5 into one line of code. This is because `group_by` permits us to group by more than one variable. Obtain one big summary table using` group_by(AgeDecade, Gender)`.
```{r}
NHANES %>% group_by(AgeDecade, Gender) %>%
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

7. For males between the ages of 40-49, compare systolic blood pressure across race as reported in the `Race1` variable. Order the resulting table from lowest to highest average systolic blood pressure.
```{r}
NHANES %>% filter(Gender == "male") %>% filter(AgeDecade == " 40-49") %>% group_by(Race1) %>%
  summarize(avg_race = mean(BPSysAve, na.rm = TRUE)) %>% arrange(avg_race)
```


# 5.10 Tibbles
n Section 5.7.3 we introduced the `group_by` function, which permits stratifying data before computing summary statistics. But where is the group information stored in the data frame?
```{r}
murders %>% group_by(region) %>% head()
```
Notice that there are no columns with this information. But, if you look closely at the output above, you see the line `A tibble: 6 x 5`. We can learn the class of the returned object using:
```{r}
murders %>% group_by(region) %>% class()
```
The `tbl`, pronounced tibble, is a special kind of data frame. The functions `group_by` and summarize always return this type of data frame. The `group_by` function returns a special kind of `tbl`, the `grouped_df.`
For consistency, the __dplyr__manipulation verbs (select, filter, mutate, and arrange) preserve the class of the input. Tibbles are the preferred format in the tidyverse and as a result tidyverse functions that produce a data frame from scratch return a tibble. 


## 5.10.1 Tibbles display better
The print method for tibbles is more readable than that of a data frame. To see this, compare the outputs of typing `murders` and the output of murders if we convert it to a tibble. We can do this using `as_tibble(murders)`.

## 5.10.2 Subsets of tibbles are tibbles
If you subset the columns of a data frame, you may get back an object that is not a data frame, such as a vector or scalar. For example:
```{r}
class(murders[,4])
```
is not a data frame. With tibbles this does not happen:
```{r}
class(as_tibble(murders)[,4])
```
This is useful in the tidyverse since functions require data frames as input.

With tibbles, if you want to access the vector that defines a column, and not get back a data frame, you need to use the accessor `$:`
```{r}
class(as_tibble(murders)$population)
```

 If we accidentally write `Population` instead of `population` this:
```{r}
murders$Population
```
returns a NULL with no warning, which can make it harder to debug. In contrast, if we try this with a tibble we get an informative warning:
```{r}
as_tibble(murders)$Population
```

## 5.10.3 Tibbles can have complex entries
Also, we can create tibbles with functions:
```{r}
tibble(id = c(1, 2, 3), func = c(mean, median, sd))
```

## 5.10.4 Tibbles can be grouped
The function `group_by` returns a special kind of tibble: a grouped tibble. 
This class stores information that lets you know which rows are in which groups. The tidyverse functions, in particular the `summarize` function, are aware of the group information.

## 5.10.5 Create a tibble using `data_frame` instead of `data.frame`
It is sometimes useful for us to create our own data frames. To create a data frame in the tibble format, you can do this by using the `data_frame` function.
```{r}
grades <- data_frame(names = c("John", "Juan", "Jean", "Yao"), 
                     exam_1 = c(95, 80, 90, 85), 
                     exam_2 = c(90, 85, 85, 90))
grades
```

One other important difference is that by default data.frame coerces characters into factors without providing a warning or message:
```{r}
grades <- data.frame(names = c("John", "Juan", "Jean", "Yao"), 
                     exam_1 = c(95, 80, 90, 85), 
                     exam_2 = c(90, 85, 85, 90))
class(grades$names)
```

To avoid this, we use the rather cumbersome argument stringsAsFactors:
```{r}
grades <- data.frame(names = c("John", "Juan", "Jean", "Yao"), 
                     exam_1 = c(95, 80, 90, 85), 
                     exam_2 = c(90, 85, 85, 90),
                     stringsAsFactors = FALSE)
class(grades$names)
```

To convert a regular data frame to a tibble, you can use the as_tibble function.
```{r}
as_tibble(grades) %>% class()
```

# 5.11 The dot operator
if we want to compute the median murder rate for states in the southern states, instead of typing:
```{r}
tab_1 <- filter(murders, region == "South")
tab_2 <- mutate(tab_1, rate = total / population * 10^5)
rates <- tab_2$rate
median(rates)
```

We can avoid defining any new intermediate objects by instead typing:
```{r}
filter(murders, region == "South") %>% 
  mutate(rate = total / population * 10^5) %>% 
  summarize(median = median(rate)) %>%
  pull(median)
```
We can do this because each of these functions takes a data frame as the first argument.

But what if the `pull` function was not available and we wanted to access `tab_2$rate`? What data frame name would we use? The answer is the dot operator.

For example to access the rate vector without the pull function we could use
```{r}
rates <-filter(murders, region == "South") %>% 
  mutate(rate = total / population * 10^5) %>% 
  .$rate
median(rates)
```

# 5.12 `do`
The `do` functions serves as a bridge between R functions such as quantile and the tidyverse. The `do` function understands grouped tibbles and always returns a data frame.

In Section 5.7.1, we noted that if we attempt to use `quantile` to obtain the min, median and max in one call, we will receive an error: 
`Error: expecting result of length one, got : 2.`
```{r}
data(heights)
heights %>% 
  filter(sex == "Female") %>%
  summarize(range = quantile(height, c(0, 0.5, 1)))
```
We can use the `do` function fix this.

First we have to write a function that fits into the tidyverse approach: that is, it receives a data frame and returns a data frame.
```{r}
my_summary <- function(dat){
  x <- quantile(dat$height, c(0, 0.5, 1))
  data_frame(min = x[1], median = x[2], max = x[3])
}
```

We can now apply the function to the heights dataset to obtain the summaries:
```{r}
heights %>% 
  group_by(sex) %>% 
  my_summary
```

We want a summary for each sex and the code returned just one summary. This is because `my_summary` is not part of the tidyverse and does not know how to handled grouped tibbles. `do` makes this connection:
```{r}
heights %>% 
  group_by(sex) %>% 
  do(my_summary(.))
```
The tibble created by `group_by` is piped to `do.` 

Within the call to `do`, the name of this tibble is `.` and we want to send it to `my_summary`. If you do not use the dot, then `my_summary` has __no argument and returns an error telling us that `argument "dat"` is missing. You can see the error by typing:
```{r}
heights %>% 
  group_by(sex) %>% 
  do(my_summary())
```


# 5.13 The `purrr` package
We learned about the `sapply` function, which permitted us to apply the same function to each element of a vector. We constructed this function:
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x)
}
```
and used sapply to compute the sum of the first n integers for several values of n like this:
```{r}
n <- 1:25
s_n <- sapply(n, compute_s_n)
```

The purrr package includes functions similar to `sapply` but that better interact with other tidyverse functions. = we can better control the output type of functions.
`sapply` can return several different object types but purr functions never do this:
= they will return objects of a specified type or return an error if this is not possible.

1) `map`, which works very similar to `sapply` but always, without exception, returns a `list`:
```{r}
library(purrr)
s_n <- map(n, compute_s_n)
class(s_n)
s_n <- map_dbl(n, compute_s_n)
class(s_n)
```
2)`map_dbl`: which always returns a vector of numeric values.

3) `map_df`: which always returns a tibble data frame.
```{r}
s_n <- map_df(n, compute_s_n)
```

```{r}
compute_s_n <- function(n){
  x <- 1:n
  data_frame(sum = sum(x))
}
s_n <- map_df(n, compute_s_n)
head(s_n)
```

# 5.14 Tidyverse conditionals
 In this section we present two dplyr functions that provide further functionality for performing conditional operations. (like `ifelse` function)

# 5.14.1 `case_when`
The `case_when` function is useful for vectorizing conditional statements. It is similar to `ifelse` but can output any number of values, as opposed to just `TRUE` or `FALSE.` 
Here is an example splitting numbers into negative, positives and 0:
```{r}
x <- c(-2, -1, 0, 1, 2)
case_when(x < 0 ~ "Negative", x > 0 ~ "Positive", TRUE ~ "Zero")
```
= define categorical variables based on existing variables.

Here is how we use case_when to do this:
```{r}
data(murders)
murders %>% 
  mutate(group = case_when(
    abb %in% c("ME", "NH", "VT", "MA", "RI", "CT") ~ "New England",
    abb %in% c("WA", "OR", "CA") ~ "West Coast",
    region == "South" ~ "South",
    TRUE ~ "other")) %>%
  group_by(group) %>%
  summarize(rate = sum(total) / sum(population) * 10^5) %>%
  arrange(rate)
```

# 5.14.2 `between`
For example to check if the elements of a vector `x` are between `a` and `b` we can type
```{r}
x >= a & x <= b
between(x, a, b)
```

# 5.15 Exercises

1. Load the `murders` dataset. Which of the following is true?
```{r}
library(dslabs)
data("murders")
class(murders)
```

A. `murders` is in tidy format and is stored in a tibble.

**B. `murders` is in tidy format and is stored in a data frame.**

C. `murders` is not in tidy format and is stored in a tibble.

D. `murders` is not in tidy format and is stored in a data frame.


2. Use `as_tibble` to covert the `murders` data table into a tibble and save it in an object called murders_tibble.
```{r}
murders_tibble <- as_tibble(murders)
class(murders_tibble)
```

3. Use the `group_by` function to convert murders into a tibble that is grouped by region.
```{r}
murders_tibble %>%
  group_by(region)
```


4. Write tidyverse code that is equivalent to this code: `exp(mean(log(murders$population)))`. Write it using the pipe so that each function is called without arguments. Use the dot operator to access the population. Hint: The code should start with `murders %>%`.
```{r}
murders_pop <- function(dat){
    x <-log(murders$population)
    data_frame(murders_pop_log = mean(x))
}
murders %>% do(murders_pop(.)) %>% exp()

murders %>% murders_pop() %>% exp()

exp(mean(log(murders$population)))
```


5. Use the `map_df` to create a data frame with three columns named `n`, `s_n`, and `s_n_2.` The first column should contain the numbers 1 through 100. The second and third columns should each contain the sum of 1 through n with n the row number.
```{r}
compute_s_n <- function(n){
  x <- 1:n
  data_frame(n = n, s_n = sum(x), s_n_2 = sum(x))
}
new_data <- map_df(n,compute_s_n)
new_data
```



