---
title: "20190918"
output: html_notebook
---

# Chapter 5 The tidyverse

지금까지는 indexing을 통해 vector를 reorder하고 subset했지만 더 복잡한 data는 data frame을 통해 분석된다. 이번에는 specific data format인 *tidy*와 *tidyverse*에 대해 알아볼 것 이다.

Install and load the tidyverse package:
```{r}
install.packages("tidyverse")
library(tidyverse)
```
These are packages included in tidyverse:
 dplyr: manipulate data frames
 purr: working with function
 ggplot2: graphing package


## 5.1 Tidy data

If each row represents one observation and columns represent the different variables available for each of these observations, it is tidy format.
The `murders`dataset of tidy data frame:
```{r}
#>        state abb region population total
#> 1    Alabama  AL  South    4779736   135
#> 2     Alaska  AK   West     710231    19
#> 3    Arizona  AZ   West    6392017   232
#> 4   Arkansas  AR  South    2915918    93
#> 5 California  CA   West   37253956  1257
#> 6   Colorado  CO   West    5029196    65
```
Each row: state
Five columns: different variable (name, abbreviation, region, population, and total murders)

In different formats:
```{r}
#>        country year fertility
#> 1      Germany 1960      2.41
#> 2  South Korea 1960      6.16
#> 3      Germany 1961      2.44
#> 4  South Korea 1961      5.99
#> 5      Germany 1962      2.47
#> 6  South Korea 1962      5.79
#> 7      Germany 1963      2.49
#> 8  South Korea 1963      5.57
#> 9      Germany 1964      2.49
#> 10 South Korea 1964      5.36
#> 11     Germany 1965      2.48
#> 12 South Korea 1965      5.16
```
Each row: one observation (country)
Three columns: variables(county, year and fertility rate)
But this was reshaped for the dslabs package.

Originally:
```{r}
#>       country 1960 1961 1962 1963 1964 1965
#> 1     Germany 2.41 2.44 2.47 2.49 2.49 2.48
#> 2 South Korea 6.16 5.99 5.79 5.57 5.36 5.16
```
Two important differences:
1) each row includes several observations
2) one of the variable, year


## 5.2 Exercises

1. Examine the built-in dataset co2. Which of the following is true:
```{r}
data("co2")
co2
```
A. co2 is tidy data: it has one year for each row.


2. Examine the built-in dataset ChickWeight. Which of the following is true:
```{r}
data("ChickWeight")
ChickWeight
```
A. ChickWeight is not tidy: each chick has more than one row.


3. Examine the built-in dataset BOD. Which of the following is true:
```{r}
data("BOD")
BOD
```
C. BOD is tidy: each row is an observation with two values (time and demand)


4. Which of the following built-in datasets is tidy (you can pick more than one):
```{r}
BJsales
EuStockMarkets
DNase
Formaldehyde
Orange
UCBAdmissions
```
C. DNase

D. Formaldehyde

E. Orange

**tidy와 다른 data set간의 차이를 명확하게 이해하지 못했습니다. 이 문제에서 다른 보기는 어떤 data인지 알고 싶습니다.**


## 5.3 Manipulating data frames

To change the data table by adding a new column, we use `mutate`.
To filter the data table to a subset of rows, we use `filter`.
To subset the data by seleting specific columns, we use `select`.


### 5.3.1 Adding a column with `mutate`

The first task is to add the murder rates to the frame using function `mutate.

-> mutate(data frame, `name = values`)

To add murder rates, we use:
```{r}
library(dslabs)
data("murders")
murders <- mutate(murders, rate = total / population * 100000)
head(murders)
```
This time, `total` and `population` are included in dplyr, and the `data(murders)` is remained.


### 5.3.2 Subsetting with `filter`

If we want to filter the data table to only show the entries for which the murder rate is lower than 0.71, use the `filter` function.

-> filter(data table, `conditional statement`)
```{r}
filter(murders,rate <= 0.71)
```


### 5.3.3 Selecting columns with `select`

If we want to view just a few, we can use the dplyr `select` function.
Select three coumns:
```{r}
new_table <- select(murders,state,region, rate)
filter(new_table, rate <= 0.71)
```
The first argument murders is an object, but state, region, and rate are variable names.
-> select(object, variable names)


## 5.4 Exercises

1. Use the function `mutate` to add a murders column named `rate` with the per 100,000 murder rate as in the example code above. Make sure you redefine `murders` as done in the example code above ( murders <- [your code]) so we can keep using this variable.
```{r}
murders <- mutate(murders, population_in_millions = population / 10^6)
murders <- mutate(murders, rate = total/population * 100000)
murders
```

2. If `rank(x)` gives you the ranks of `x` from lowest to highest, `rank(-x)` gives you the ranks from highest to lowest. Use the function `mutate` to add a column `rank` containing the rank, from highest to lowest murder rate. Make sure you redefine `murders` so we can keep using this variable.
```{r}
murders <- mutate(murders, rank = rank(murders$rate))
murders
```

3. Use `select` to show the state names and abbreviations in `murders`. Do not redefine `murders`, just show the results.
```{r}
select(murders, state, abb)
```


4. The dplyr function `filter` is used to choose specific rows of the data frame to keep. Unlike `select` which is for columns, `filter` is for rows. For example, you can show just the New York row like this:
```{r}
filter(murders, state == "New York")
```

Use `filter` to show the top 5 states with the highest murder rates. After we add murder rate and rank, do not change the murders dataset, just show the result. Remember that you can filter based on the `rank` column.
```{r}
filter(murders, rank <= 5)
```

5. We can remove rows using the `!=` operator. For example, to remove Florida, we would do this:
```{r}
no_florida <- filter(murders, state != "Florida")
```
Create a new data frame called `no_south` that removes states from the South region. How many states are in this category? You can use the function `nrow` for this.
```{r}
no_south <- filter(murders, region != "South")
nrow(no_south)
```

6. We can also use `%in%` to filter with dplyr. You can therefore see the data from New York and Texas like this:
```{r}
filter(murders, state %in% c("New York", "Texas"))
```

Create a new data frame called `murders_nw` with only the states from the Northeast and the West. How many states are in this category?
```{r}
murders_nw <- filter(murders, region %in% c("Northeast", "West"))
murders_nw
```


7. Suppose you want to live in the Northeast or West and want the murder rate to be less than 1. We want to see the data for the states satisfying these options. Note that you can use logical operators with `filter`. Here is an example in which we filter to keep only small states in the Northeast region.
```{r}
filter(murders, population < 5000000 & region == "Northeast")
```
Make sure `murders` has been defined with `rate` and `rank` and still has all states. Create a table called `my_states` that contains rows for states satisfying both the conditions: it is in the Northeast or West and the murder rate is less than 1. Use `select` to show only the state name, the rate and the rank.
```{r}
my_states <- filter(murders, rate < 1 & region %in% c("Northeast", "West"))
select(my_states, state, rate, rank)
```


## 5.5 The pipe: `%>%`
pipe operator,`%>%`를 이용해 series of operation을 진행할 수 있다. 
Original data -> select -> filter:
```{r}
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
```

The pipe sends the result of the left side to be the first argument of the fnction on the right side, like this.
```{r}
16 %>% sqrt()
16 %>% sqrt() %>% log2()
```
The above statement is equivalent to `log2(sqrt(16)).

```{r}
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
```
murders: first argument of the select function
new data grame: first argument of the filter function


## 5.6 Exercises

1. The pipe `%>%` can be used to perform operations sequentially without having to define intermediate objects. Start by redefining murder to include rate and rank.
```{r}
murders <- mutate(murders, rate =  total / population * 100000, rank = rank(-rate))
```
In the solution to the previous exercise, we did the following:
```{r}
my_states <- filter(murders, region %in% c("Northeast", "West") & rate < 1)

select(my_states, state, rate, rank)
```
The pipe` %>%` permits us to perform both operations sequentially without having to define an intermediate variable `my_states`. We therefore could have mutated and selected in the same line like this:
```{r}
mutate(murders, rate =  total / population * 100000, rank = rank(-rate)) %>%
  select(state, rate, rank)
```
Notice that `select` no longer has a data frame as the first argument. The first argument is assumed to be the result of the operation conducted right before the `%>%`.

Repeat the previous exercise, but now instead of creating a new object, show the result and only include the state, rate, and rank columns. Use a pipe `%>%` to do this in just one line.
```{r}
filter(murders, region %in% c("Northeast", "West") & rate < 1) %>% select(state, rate, rank)
```

2. Reset `murders` to the original table by using `data(murders)`. Use a pipe to create a new data frame called` my_states` that considers only states in the Northeast or West which have a murder rate lower than 1, and contains only the state, rate and rank columns. The pipe should also have four components separated by three `%>%`. The code should look something like this:
```{r}
data("murders")

my_states <- murders %>% mutate(rate = total/population*100000, rank = rank(rate)) %>% filter(region %in% c("Northeast", "West") & rate < 1) %>% select(state, rate, rank)
my_states
```


